---
title: "LAGOS Spatial Analysis"
author: "Amanda Hastings"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(RApiSerialize)
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
#Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus


# load('lake_centers.Rdata')

```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
#Show a map outline of Iowa and Illinois 

i_states <- states %>% 
  filter(name %in% c('Iowa','Illinois')) %>% 
  st_transform(2163)

mapview(i_states)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

```{r}
#Subset LAGOS data

istate_lakes <- spatial_lakes[i_states,]

nrow(minnesota_lakes)-nrow(istate_lakes)

```

There are 16446 sites in Illinois and Iowa combined. Meanwhile, there are 29038 Minnesota sites, exceeding the number of sites in Illinois/Iowa by 12572 sites. 


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
iowa <- states %>%
  filter(name == 'Iowa') %>%
  st_transform(2163)

iowa_lakes <- spatial_lakes[iowa,]

ggplot(iowa_lakes, aes(x=lake_area_ha))+ 
  geom_histogram(bins=40)+ 
  scale_x_log10()+
  labs(x= "Lake size (hectares)", y= "Frequency", title='Iowa Lakes')

ggplot(minnesota_lakes, aes(x=lake_area_ha))+ 
  geom_histogram(bins=40)+ 
  scale_x_log10()+
  labs(x= "Lake size (hectares)", y= "Frequency", title='Minnesota Lakes')
```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them by lake area in hectares. 

```{r}
istate_lakes %>%
  arrange(-lake_area_ha) %>%
  slice(1:1000) %>%
  mapview(., zcol = 'lake_area_ha')
```


## 5) What other data sources might we use to understand how reservoirs and natural lakes vary in size in these three states? 


