---
title: "Lake Water Quality Analysis"
author: "Amanda Hastings"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(lubridate) #For dealing with date and time
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
#lagosne_get(dest_folder = LAGOSNE:::lagos_path(),overwrite=T)

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
#names(nutr)
```

### Subset columns nutr to only keep key info that we want


```{r}
clarity_only <- nutr %>%
  select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% ymd(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observations did we lose?
# nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)


```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                            distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')


```

### Mean Chl_a map

```{r}
### Take the mean chl_a and secchi by lake

mean_values_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,mean_values_200,
                          by='lagoslakeid') 

#Make a map
mapview(mean_spatial,zcol='log10_mean_chl')
```


# Class work

## 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for 
sites with at least 200 observations?

- Here, I just want a plot of chla vs secchi for all sites 

```{r}
#Plot secchi disk depth by cholorphyll a 
library(ggthemes)
ggplot(chla_secchi_200, aes(x = secchi, y = chla)) +
  geom_point(shape = 1, color="#48680E")+ 
  theme_few()+
  labs(y="Chlorophyll A", x="Secchi Disk Depth")+ 
  xlim(0,16)

```


```{r include=FALSE}
cor.test(chla_secchi_200$chla,chla_secchi_200$secchi)
```

There is a negative correlation between secchi disk depth and chlorophyll a levels, where shallower secchi disk depths are associated with higher chlorophyll a levels. 


## Why might this be the case? 

High cholorphyll a concentrations are associated with high biomass of photosynthetic organisms, such as algae. Meanwhile, secchi disks are used to assess water turbidity, where the depth of disk disappearance indicates the transparency of the water(epa.gov). Where 

[epa.gov] 

Tilzer, M. M. (1988). Secchi disk - chlorophyll relationships in a lake with highly variable phytoplankton biomass. Hydrobiologia, 162(2), 163–171. https://doi.org/10.1007/BF00014539

## 2) What states have the most data? 

### 2a) First you will need to make a lagos spatial dataset that has the total 
number of counts per site.

```{r}
#Create lagos spatial dataset with total number of counts per site 

site_counts <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n())

#Add geographic information to site counts data 
geo_counts <- inner_join(site_counts,lake_centers, by= 'lagoslakeid')%>% 
  select(lagoslakeid,nhd_long,nhd_lat, count, secchi, chla)

#Create spatial object for lagos dataset 
spatial_counts <- st_as_sf(geo_counts,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

###Option 1: Use state_zoneid to join data 
#Parse out lagos states data to be joined to original lagos spatial dataset  
  
# states_lagos <- lagos$state %>%
#   select(-state_pct_in_nwi, -state_ha_in_nwi,-state_ha)
# 
# #Join lagos state data and original lagos spatial data by state_zoneid field 
# 
# stateid_counts <- full_join(site_counts, states_lagos, by="state_zoneid")
# 
# #Calculate the sum of counts per state 
# 
# state_sums <- stateid_counts %>%
#   group_by(state_name) %>%
#   summarize(sum_counts = sum(count))

  
```


### 2b) Second, you will need to join this point dataset to the us_boundaries data.

```{r}
###Option 2: Use a spatial join to combine point dataset and us_boundaries data 
#Parse out state boundaries from us_boundaries data   

states <- us_states()


#Use spatial join to overlay point dataset (original lagos spatial dataset) and state boundaries 
overlay <- st_join(spatial_counts,states)

```


### 2c) Then you will want to group by state and sum all the observations in 
that state and arrange that data from most to least total observations per state. 

```{r}
###Option 1:
#Join us_states data to lagos dataset by state_name
#Arrange summed counts in descending order 

# states_join <- inner_join(states,state_sums,by='state_name') %>% 
#   arrange(desc(sum_counts)) %>% 
#   select(state_name,sum_counts)

#Plot summed counts by state

# mapview(states_join, zcol='sum_counts', layer.name = 'Observations per state',canvas=TRUE)
# 
# head(states_join)

```


```{r}
# #Option 2: 
# Group by state, sum number of observations by state
summed_overlay <- overlay %>%
  group_by(state_name) %>%
  summarize(sum_count = sum(count)) %>%
  arrange(desc(sum_count))

# head(summed_overlay)
  
  
```


The top five states with the most data include: Minnesota, Michigan, Wisconsin, New York, and Illinois. 


##3 Is there a spatial pattern in Secchi disk depth for lakes with at least 200 
observations?

```{r}
## Your code here

secchi200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  summarize(mean_secchi = mean(secchi), count = n()) %>%
  filter(count > 200)
  

secchi200_locus <- inner_join(secchi200, lake_centers, by='lagoslakeid') %>%
  select(lagoslakeid, nhd_long, nhd_lat, mean_secchi)

spatial_secchi200 <- st_as_sf(secchi200_locus,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

mapview(spatial_secchi200, zcol='mean_secchi', layer.name='Secchi Disk Depth', canvas=TRUE)

```

I do not see a strong spatial pattern in secchi disk depth in lakes with at least 200 observations. 
